---
pagetitle: "Trial plan"
execute:
  freeze: false
  echo: false
format:
  html:
    page-layout: custom
    toc: false
---

<!--
PROTOTYPE. A field-trial take on the catalogue, kept separate from
dtable.qmd so the two can be compared. Styles are inline for the same
reason: deleting this file removes the whole experiment.

The page is a full-viewport plan drawn as a line diagram, the way a
design package or a methods paper would draw a trial. Each package is
a plot; blocks are organisations; fill is a single green ramp for
"vigour" (time since last commit). Click a plot for its entry.
-->

```{r}
#| include: false
library(here)
library(readr)
library(dplyr)
library(htmltools)

df <- read_csv(here("data/processed/submissions_cleaned.csv"), show_col_types = FALSE)

metrics_path <- here("data/processed/package_metrics.csv")
if (file.exists(metrics_path)) {
  df <- left_join(df, read_csv(metrics_path, show_col_types = FALSE), by = "Software website")
}
for (col in c("github_stars", "github_followers", "first_commit_date", "last_commit_date",
              "contributors", "cran_package", "cran_downloads_last_month", "cran_downloads_total")) {
  if (!col %in% names(df)) df[[col]] <- NA
}

esc <- function(x) htmltools::htmlEscape(ifelse(is.na(x), "", as.character(x)))
fmt_n <- function(x) if (is.na(x)) "—" else format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
`%||%` <- function(a, b) if (is.null(a) || is.na(a)) b else a

derive_name <- function(url) {
  if (is.na(url) || !nzchar(url)) return(NA_character_)
  u <- sub("/+$", "", sub("^https?://", "", url))
  parts <- strsplit(u, "/")[[1]]
  host <- parts[1]
  if (grepl("github\\.io$", host) && length(parts) >= 2) return(parts[2])
  if (identical(host, "github.com") && length(parts) >= 3) return(parts[3])
  if (length(parts) >= 2) return(parts[length(parts)])
  sub("\\..*$", "", host)
}
split_tags <- function(x) {
  if (is.na(x) || !nzchar(x)) return(character(0))
  out <- trimws(unlist(strsplit(x, ",")))
  out[nzchar(out)]
}

today <- Sys.Date()

vigour <- function(last) {
  if (is.na(last)) return(list(lvl = 0, label = "Unscored", ago = ""))
  d <- as.numeric(today - as.Date(last))
  ago <- if (d < 1) "today" else if (d < 2) "yesterday" else if (d < 60) sprintf("%d days ago", d)
         else if (d < 700) sprintf("%d months ago", round(d / 30.4)) else sprintf("%d years ago", round(d / 365))
  if (d <= 90)       list(lvl = 3, label = "Active",     ago = ago)
  else if (d <= 365) list(lvl = 2, label = "Maintained", ago = ago)
  else if (d <= 730) list(lvl = 1, label = "Quiet",      ago = ago)
  else               list(lvl = 0, label = "Dormant",    ago = ago)
}

df <- df |>
  mutate(name  = vapply(`Software website`, derive_name, character(1)),
         block = coalesce(Organisation, "Unaffiliated")) |>
  arrange(block, desc(coalesce(as.Date(last_commit_date), as.Date("1900-01-01"))))

blocks <- split(df, df$block)
blocks <- blocks[order(-vapply(blocks, nrow, integer(1)), names(blocks))]
all_keywords <- sort(unique(unlist(lapply(df$`Code keywords`, split_tags))))

# ---- Geometry (plan units) -------------------------------------------------
# Sizes only. The page script lays the blocks out to suit the screen: side
# by side with roughly square blocks on wide screens, stacked in a single
# column on phones. Short blocks get buffer plots so each is a rectangle.
PW <- 14; PH <- 8.5        # plot footprint, laid out landscape
CG <- 1; RG <- 2           # column gap; alley between ranges
BG <- 10                   # minimum gap between blocks
PAD <- 8

# Plot IDs are fixed here (block letter + order within the block) so they
# stay the same whatever the screen layout.
plots_tbl <- bind_rows(lapply(seq_along(blocks), function(bi) {
  b <- blocks[[bi]]
  b$letter <- LETTERS[bi]
  b$idx    <- seq_len(nrow(b))
  b$id     <- sprintf("%s%02d", LETTERS[bi], b$idx)
  b
}))

# ---- SVG ---------------------------------------------------------------------
svg_text <- function(x, y, s, cls, extra = "") sprintf('<text class="%s" x="%.2f" y="%.2f"%s>%s</text>', cls, x, y, extra, esc(s))

# ---- Crop glyphs ------------------------------------------------------------
# Each plot is sown with one plant per contributor, drawn at the growth
# stage that matches its activity: in ear (active), leafy (maintained),
# seedling (quiet) or stubble (dormant). Glyphs are drawn at the origin,
# growing up the negative y axis, in plan units.
glyph <- function(lvl) switch(as.character(lvl),
  "3" = paste0(
    '<path class="stem" d="M0 0 Q0.06 -1.2 0 -2.1"/>',
    '<path class="leaf" d="M0 -0.7 Q-0.55 -1 -0.9 -0.75"/>',
    '<path class="leaf" d="M0 -1.3 Q0.55 -1.6 0.85 -1.35"/>',
    paste0(vapply(0:4, function(j) sprintf(
      '<ellipse class="grain" cx="%.2f" cy="%.2f" rx="0.11" ry="0.19" transform="rotate(%d %.2f %.2f)"/>',
      if (j %% 2) 0.11 else -0.11, -2.25 - j * 0.17, if (j %% 2) 22 else -22,
      if (j %% 2) 0.11 else -0.11, -2.25 - j * 0.17), character(1)), collapse = ""),
    '<path class="awn" d="M0 -2.95 L-0.08 -3.4 M0.11 -2.75 L0.36 -3.2 M-0.11 -2.6 L-0.38 -3.05"/>'),
  "2" = paste0(
    '<path class="stem" d="M0 0 Q0.05 -1 0 -2"/>',
    '<path class="leaf" d="M0 -0.55 Q-0.5 -0.85 -0.85 -0.65"/>',
    '<path class="leaf" d="M0 -1.05 Q0.5 -1.35 0.8 -1.15"/>',
    '<path class="leaf" d="M0 -1.55 Q-0.3 -1.95 -0.55 -2.1"/>'),
  "1" = paste0(
    '<path class="stem" d="M0 0 L0 -0.55"/>',
    '<path class="leaf" d="M0 -0.45 Q-0.15 -0.85 -0.5 -0.95"/>',
    '<path class="leaf" d="M0 -0.5 Q0.2 -0.95 0.45 -1.05"/>'),
  '<path class="stubble" d="M-0.12 0 L-0.18 -0.45 M0.12 0 L0.16 -0.4"/>')

# Deterministic wobble so the stand looks hand-sown but is identical on
# every render.
wobble <- function(i, seed) sin(i * 12.9898 + seed * 78.233)

# A row of plants across [x0, x1] on a baseline at y. `delay0` staggers the
# grow-in animation across the field.
crop_row <- function(lvl, n, x0, x1, y, seed = 1, scale = 1, delay0 = 0) {
  n  <- max(1, min(n, 14))
  xs <- x0 + (x1 - x0) * (seq_len(n) - 0.5) / n
  paste0(vapply(seq_len(n), function(i) {
    h    <- scale * (1 + 0.08 * wobble(i, seed))
    lean <- 4 * wobble(i + 7, seed)
    sprintf('<g transform="translate(%.2f %.2f)"><g class="plant" style="--d:%.2fs"><g class="sway" style="--s:%.2fs"><g class="g-v%d" transform="rotate(%.1f) scale(%.2f)">%s</g></g></g></g>',
            xs[i], y, delay0 + i * 0.03, 0.15 * abs(wobble(i, seed + 3)), lvl, lean, h, glyph(lvl))
  }, character(1)), collapse = "")
}

plot_svg <- function(r) {
  v  <- vigour(r$last_commit_date)
  ck <- split_tags(r$`Code keywords`); tk <- split_tags(r$`Topic keywords`)
  haystack <- tolower(paste(r$name, r$`Code maintainer`, r$Organisation, r$Description,
                            paste(ck, collapse = " "), paste(tk, collapse = " "),
                            r$Language, r$License, r$`Code type`))
  # Shrink long names so they always fit across the plot.
  name_size <- min(1.6, (PW - 1.2) / (0.62 * max(1, nchar(r$name %||% ""))))
  by_size   <- min(0.9, (PW - 1.2) / (0.5 * max(1, nchar(r$`Code maintainer` %||% ""))))
  n_plants  <- if (is.na(r$contributors)) 3 else r$contributors
  base_y    <- PH - 0.55
  seed      <- sum(utf8ToInt(r$id))
  sprintf('
  <g class="plot is-v%d" tabindex="0" role="button" data-id="%s" data-search="%s" data-keywords="%s"
     data-status="%s" data-cran="%d">
    <title>%s — %s</title>
    <rect class="plot-fill" width="%g" height="%g"/>
    <line class="soil" x1="0.5" x2="%.2f" y1="%.2f" y2="%.2f"/>
    <g class="crop">%s</g>
    %s
    %s
    <text class="plot-name" x="0.6" y="3" font-size="%.2f">%s</text>
    <text class="plot-by" x="0.6" y="4.15" font-size="%.2f">%s</text>
  </g>',
    v$lvl, r$id, esc(haystack), esc(tolower(paste(ck, collapse = "|"))),
    tolower(v$label), as.integer(!is.na(r$cran_package)),
    esc(r$name), esc(r$`Code maintainer`),
    PW, PH,
    PW - 0.5, base_y, base_y,
    crop_row(v$lvl, n_plants, 0.6, PW - 0.6, base_y, seed = seed, scale = 0.9,
             delay0 = 0.05 * (match(r$letter, LETTERS) - 1) + 0.02 * r$idx),
    svg_text(0.55, 1.35, r$id, "plot-id"),
    if (!is.na(r$cran_package)) sprintf('<circle class="plot-cran" cx="%.2f" cy="1.0" r="0.38"><title>On CRAN</title></circle>', PW - 0.95) else "",
    name_size, esc(r$name),
    by_size, esc(r$`Code maintainer`))
}

block_svg <- function(bi) {
  b <- plots_tbl[plots_tbl$letter == LETTERS[bi], ]
  paste0(
    sprintf('<g class="block" data-letter="%s">', LETTERS[bi]),
    sprintf('<text class="block-cap" data-name="%s"><tspan class="block-letter">Block %s</tspan></text>',
            esc(names(blocks)[bi]), LETTERS[bi]),
    '<g class="block-extras"></g>',
    paste0(vapply(seq_len(nrow(b)), function(i) plot_svg(b[i, ]), character(1)), collapse = ""),
    '</g>')
}
field_svg <- paste0(vapply(seq_along(blocks), block_svg, character(1)), collapse = "\n")

# ---- Entry panel content, one template per plot -----------------------------
row <- function(k, v, hint = NULL) sprintf('<div class="e-row"%s><dt>%s</dt><dd>%s</dd></div>',
                                            if (is.null(hint)) "" else sprintf(' title="%s"', esc(hint)), esc(k), v)
detail_html <- function(r) {
  v <- vigour(r$last_commit_date)
  ck <- split_tags(r$`Code keywords`); tk <- split_tags(r$`Topic keywords`)
  tags <- function(x) if (length(x)) paste0('<p class="e-tags">', paste0(esc(x), collapse = '<span class="dot">·</span>'), '</p>') else '<p class="e-tags">—</p>'
  sprintf('
  <template data-for="%s">
    <svg class="e-crop" viewBox="0 0 26 4.2" aria-hidden="true"><line class="soil" x1="0" x2="26" y1="3.9" y2="3.9"/>%s</svg>
    <p class="e-kicker">Plot %s · Block %s · <span class="e-pos"></span></p>
    <h2 class="e-name">%s</h2>
    <p class="e-by">%s · %s</p>
    <p class="e-status">%s%s</p>
    <p class="e-desc">%s</p>
    <dl class="e-list">%s%s%s%s%s%s%s%s%s%s</dl>
    <h3 class="e-h">Code keywords</h3>%s
    <h3 class="e-h">Topics</h3>%s
    <p class="e-links"><a href="%s" target="_blank" rel="noopener">Repository ↗</a><a href="mailto:%s">Email maintainer</a></p>
  </template>',
    r$id,
    crop_row(v$lvl, if (is.na(r$contributors)) 3 else r$contributors, 1.2, 24.8, 3.9,
             seed = sum(utf8ToInt(r$id)), scale = 1),
    r$id, r$letter,
    esc(r$name %||% r$`Code type`), esc(r$`Code maintainer`), esc(r$Organisation),
    v$label, if (nzchar(v$ago)) sprintf('<span> · last commit %s</span>', v$ago) else "",
    esc(r$Description),
    row("Language", esc(r$Language)), row("Licence", esc(r$License)), row("Type", esc(r$`Code type`)),
    row("CRAN", if (is.na(r$cran_package)) "—" else esc(r$cran_package)),
    row("First commit", if (is.na(r$first_commit_date)) "—" else format(as.Date(r$first_commit_date), "%b %Y")),
    row("Last commit", if (is.na(r$last_commit_date)) "—" else format(as.Date(r$last_commit_date), "%d %b %Y")),
    row("GitHub stars", fmt_n(r$github_stars)),
    row("Contributors", fmt_n(r$contributors)),
    row("CRAN downloads, last month", fmt_n(r$cran_downloads_last_month), "cranlogs, last 30 days"),
    row("CRAN downloads, all time", fmt_n(r$cran_downloads_total), "cranlogs, since Oct 2012"),
    tags(ck), tags(tk),
    esc(r$`Software website`), esc(r$Email))
}
templates <- paste0(vapply(seq_len(nrow(plots_tbl)), function(i) detail_html(plots_tbl[i, ]), character(1)), collapse = "\n")

legend_html <- paste0(vapply(list(
    list(3, "Active", "&lt;3 mo"), list(2, "Maintained", "&lt;1 yr"),
    list(1, "Quiet", "&lt;2 yr"), list(0, "Dormant", "")),
  function(x) sprintf('<span><svg class="lg-crop" viewBox="-1.1 -3.5 2.2 3.6" aria-hidden="true"><g class="g-v%d">%s</g><line class="soil" x1="-1" x2="1" y1="0" y2="0"/></svg>%s%s</span>',
                      x[[1]], glyph(x[[1]]), x[[2]], if (nzchar(x[[3]])) sprintf(' <small>%s</small>', x[[3]]) else ""),
  character(1)), collapse = "\n      ")

# Facet chips reuse the site's .chip style (see assets/aagi.scss) so the
# filters read the same as on the database page. Each carries a live count.
chip <- function(group, key, label, icon = "") sprintf(
  '<button type="button" class="chip facet-chip" data-group="%s" data-key="%s" aria-pressed="false">%s<span class="chip-label">%s</span><span class="chip-n" aria-hidden="true"></span></button>',
  group, esc(key), icon, esc(label))

kw_chips <- paste(vapply(all_keywords, function(k) chip("kw", tolower(k), k), character(1)), collapse = "")
status_chips <- paste(vapply(list(c(3, "Active"), c(2, "Maintained"), c(1, "Quiet"), c(0, "Dormant")), function(x)
  chip("status", tolower(x[2]), x[2],
       sprintf('<svg class="chip-crop" viewBox="-1.1 -3.5 2.2 3.6" aria-hidden="true"><g class="g-v%s">%s</g></svg>', x[1], glyph(x[1]))),
  character(1)), collapse = "")
cran_chip <- chip("cran", "1", "On CRAN", '<i class="chip-dot" aria-hidden="true"></i>')
```

```{r}
#| output: asis
cat('```{=html}\n')
cat(sprintf('
<div class="plan" id="plan">
  <header class="plan-bar">
    <div class="plan-head">
      <div class="plan-title">
        <h1>Trial plan</h1>
        <p role="status"><strong id="n-match">%d</strong> of %d plots · %d blocks · %s</p>
      </div>
      <div class="plan-tools">
        <label class="visually-hidden" for="pl-search">Search plots</label>
        <input id="pl-search" class="db-search plan-search" type="search" autocomplete="off"
               placeholder="Search packages or people…">
        <button type="button" class="filter-btn" id="pl-filter-btn" aria-expanded="false" aria-controls="pl-filters">
          <svg class="filter-icon" viewBox="0 0 16 16" aria-hidden="true"><path d="M2 4h12M4 8h8M6.5 12h3"/></svg>
          <span>Filters</span><span class="filter-badge" id="pl-filter-n" hidden></span>
        </button>
      </div>
    </div>

    <div class="filter-panel" id="pl-filters" role="dialog" aria-labelledby="fp-title" hidden>
      <div class="fp-head">
        <h2 class="fp-title" id="fp-title">Filters</h2>
        <button type="button" class="fp-close" id="pl-filters-close" aria-label="Close filters">×</button>
      </div>
      <div class="fp-body">
        <section class="facet" role="group" aria-labelledby="f-kw"><h3 class="facet-label" id="f-kw">Keyword</h3><div class="facet-chips">%s</div></section>
        <section class="facet" role="group" aria-labelledby="f-st"><h3 class="facet-label" id="f-st">Status</h3><div class="facet-chips">%s</div></section>
        <section class="facet" role="group" aria-labelledby="f-av"><h3 class="facet-label" id="f-av">Availability</h3><div class="facet-chips">%s</div></section>
      </div>
      <div class="fp-foot">
        <button type="button" class="chip chip-clear" id="pl-clear" hidden>Clear all</button>
        <button type="button" class="fp-show" id="pl-show">Show plots</button>
      </div>
    </div>
  </header>

  <div class="plan-stage">
    <div class="plan-active" id="pl-active" aria-label="Active filters" hidden></div>
    <svg class="plan-svg" id="plan-svg" data-pw="%g" data-ph="%g" data-cg="%g" data-rg="%g" data-bg="%g" data-pad="%g" xmlns="http://www.w3.org/2000/svg" aria-label="Trial plan of shared packages">
      <g id="world">
        <rect class="ground" x="0" y="0" width="1" height="1"/>
        %s
      </g>
    </svg>

    <div class="legend" id="legend">
      <button type="button" class="legend-toggle" aria-expanded="false" aria-controls="legend-items">Key</button>
      <div class="legend-items" id="legend-items" aria-label="Key">
        %s
        <span><i class="sw is-buffer"></i>Buffer</span>
        <span><i class="sw is-cran"></i>On CRAN</span>
        <span class="legend-note">One plant per contributor</span>
      </div>
    </div>

    <div class="plan-empty" id="pl-empty" hidden>
      <div class="plan-empty-card" role="status">
        <p class="plan-empty-title">No plots match</p>
        <p class="plan-empty-hint">Try fewer filters or a broader search.</p>
        <button type="button" class="chip chip-clear" id="pl-empty-clear">Clear all</button>
      </div>
    </div>

    <div class="zoom" role="group" aria-label="Zoom">
      <button type="button" id="z-out" aria-label="Zoom out">−</button>
      <button type="button" id="z-fit">Fit</button>
      <button type="button" id="z-in" aria-label="Zoom in">+</button>
    </div>

    <aside class="entry" id="entry" aria-label="Plot entry" aria-hidden="true">
      <button type="button" class="e-close" id="e-close" aria-label="Close">×</button>
      <div class="e-body" id="e-body"></div>
    </aside>
  </div>

  <div hidden id="e-templates">%s</div>
</div>
',
  nrow(df), nrow(df), length(blocks), format(today, "%d %b %Y"),
  kw_chips, status_chips, cran_chip, PW, PH, CG, RG, BG, PAD, field_svg, legend_html, templates))
cat('\n```\n')
```

```{=html}
<style>
body:has(#plan) { overflow: hidden; }
body:has(#plan) #quarto-content, body:has(#plan) main { padding: 0 !important; margin: 0 !important; max-width: none !important; }

.plan {
  --green: #00573c; --green-deep: #003d2a; --sage: #8cc0ae; --mint: #e4efe9;
  --ink: #1c1c1c; --muted: #5d6560; --rule: #dde2de; --surface: #f7f9f8;
  /* plot tints by activity; the crops carry the colour */
  --v3: #eef6f1; --v2: #f3f8f5; --v1: #f8fbf9; --v0: #fbfaf7;
  /* crop inks by growth stage */
  --c3: #1f6b45; --c3-ear: #4f8a3c; --c2: #4f9a6c; --c1: #8cc0ae; --c0: #bcae92;
  --soil: #ddd3c1;
  --mono: "JetBrains Mono", ui-monospace, Menlo, monospace;
  --head: "Poppins", sans-serif;
  display: flex; flex-direction: column;
  height: calc(100dvh - var(--nav-h, 58px));
  background: #fff; color: var(--ink);
  font-family: "Source Sans 3", sans-serif;
}

/* ---- top bar ---- */
.plan-bar { position: relative; padding: .9rem 2rem; border-bottom: 1px solid var(--rule); }
.plan-head { display: flex; align-items: center; justify-content: space-between; gap: 1.5rem; }
.plan-title h1 { margin: 0; font: 600 1.35rem var(--head); letter-spacing: -.01em; }
.plan-title p { margin: .1rem 0 0; font-size: .85rem; color: var(--muted); }
.plan-title p strong { font-family: var(--mono); font-weight: 500; color: var(--ink); }
/* .db-search and .chip come from assets/aagi.scss; only layout tweaks here */
.plan-tools { display: flex; align-items: center; gap: .6rem; flex: 0 1 32rem; min-width: 0; }
.plan-search { flex: 1 1 auto; min-width: 0; font-size: .95rem; padding: .5rem 1rem; }

/* Filters button: same pill language as the search box */
.filter-btn {
  display: inline-flex; align-items: center; gap: .45rem; flex: none;
  height: 2.5rem; padding: 0 1rem;
  font: inherit; font-size: .92rem; color: var(--ink); white-space: nowrap;
  background: #fff; border: 1.5px solid var(--rule); border-radius: 999px; cursor: pointer;
  transition: border-color .14s, background .14s, box-shadow .14s;
}
.filter-btn:hover { border-color: var(--sage); background: var(--surface); }
.filter-btn:focus-visible, .filter-btn[aria-expanded="true"] {
  outline: none; border-color: var(--sage); background: var(--surface); box-shadow: 0 0 0 3px rgba(140,192,174,.38);
}
.filter-btn.is-active { border-color: var(--green); color: var(--green); font-weight: 600; }
.filter-icon { width: 1rem; height: 1rem; fill: none; stroke: currentColor; stroke-width: 1.7; stroke-linecap: round; }
.filter-badge {
  display: inline-flex; align-items: center; justify-content: center; min-width: 1.3rem; height: 1.3rem; padding: 0 .35rem;
  border-radius: 999px; background: var(--green); color: #fff; font: 600 .72rem var(--mono);
}
.filter-badge[hidden] { display: none; }

/* Filters panel: drops from the button over the map */
.filter-panel {
  position: absolute; top: calc(100% - .4rem); right: 2rem; z-index: 6;
  display: flex; flex-direction: column; width: min(30rem, calc(100vw - 2rem)); max-height: min(70vh, 34rem);
  background: #fff; border: 1px solid var(--rule); border-radius: .6rem; box-shadow: 0 14px 40px rgba(28,28,28,.12);
}
.filter-panel[hidden] { display: none; }
.fp-head { display: flex; align-items: center; justify-content: space-between; padding: .7rem .7rem .2rem 1.15rem; }
.fp-title { margin: 0; font: 600 .98rem var(--head); }
/* Quarto underlines every h2; these are labels, not section headings */
.fp-title, .e-name { border-bottom: 0; padding-bottom: 0; }
.fp-close {
  width: 2.25rem; height: 2.25rem; border: 0; border-radius: 50%; background: none;
  font-size: 1.3rem; line-height: 1; color: var(--muted); cursor: pointer;
}
.fp-close:hover { background: var(--surface); color: var(--ink); }
.fp-body { overflow-y: auto; padding: .35rem 1.15rem .4rem; }
.facet { margin: 0 0 1rem; }
.facet-label { margin: 0 0 .45rem; font: 600 .66rem var(--head); text-transform: uppercase; letter-spacing: .08em; color: var(--muted); }
.facet-chips { display: flex; flex-wrap: wrap; gap: .35rem; }
.fp-foot { display: flex; align-items: center; gap: .75rem; padding: .7rem 1.15rem; border-top: 1px solid var(--rule); }
.fp-foot .chip-clear { margin: 0; }
.fp-foot .chip-clear[hidden] { display: none; }
.fp-show {
  margin-left: auto; padding: .5rem 1.15rem; border: 0; border-radius: 999px;
  background: var(--green); color: #fff; font: inherit; font-size: .9rem; font-weight: 600; cursor: pointer;
}
.fp-show:hover { background: var(--green-deep); }
.fp-show:disabled { background: #b9c3be; cursor: default; }
.fp-show:focus-visible { outline: none; box-shadow: 0 0 0 3px rgba(140,192,174,.6); }

/* active filters, floating over the top of the map */
.plan-active {
  position: absolute; top: .75rem; left: 1rem; right: 1rem; z-index: 2;
  display: flex; flex-wrap: wrap; align-items: center; gap: .35rem; pointer-events: none;
}
.plan-active[hidden] { display: none; }
.plan-active > * { pointer-events: auto; }
.chip.active-pill {
  display: inline-flex; align-items: center; gap: .4rem; font-size: .8rem; padding: .22rem .55rem .22rem .75rem;
  background: var(--mint); border-color: var(--sage); color: var(--green-deep);
}
.chip.active-pill:hover { background: #d6e8de; border-color: var(--green); }
.active-pill .x { font-size: 1rem; line-height: 1; opacity: .65; }
.active-clear {
  padding: .2rem .4rem; border: 0; background: rgba(255,255,255,.85); border-radius: .25rem;
  font: inherit; font-size: .8rem; color: var(--muted); text-decoration: underline; cursor: pointer;
}
.active-clear:hover { color: var(--ink); }

.facet-chip { display: inline-flex; align-items: center; gap: .35rem; }
.chip-n { font-family: var(--mono); font-size: .72rem; color: var(--muted); min-width: .6rem; text-align: center; }
.facet-chip[aria-pressed="true"] .chip-n { color: rgba(255,255,255,.75); }
.facet-chip:disabled { opacity: .4; cursor: default; }
.facet-chip:disabled:hover { border-color: var(--rule); background: #fff; }
.chip-crop { width: .7rem; height: 1.15rem; margin: -.2rem 0 -.1rem -.15rem; overflow: visible; }
.facet-chip[aria-pressed="true"] .chip-crop :is(.stem, .leaf, .awn, .stubble) { stroke: #fff; }
.facet-chip[aria-pressed="true"] .chip-crop .grain { fill: #fff; }
.chip-dot { width: .45rem; height: .45rem; border-radius: 50%; background: var(--green); }
.facet-chip[aria-pressed="true"] .chip-dot { background: #fff; }

/* empty state, centred over the map */
.plan-empty {
  position: absolute; inset: 0; z-index: 1; display: flex; align-items: center; justify-content: center;
  pointer-events: none;
}
.plan-empty[hidden] { display: none; }
.plan-empty-card {
  pointer-events: auto; display: flex; flex-direction: column; align-items: center; gap: .3rem;
  padding: 1.25rem 1.75rem; text-align: center;
  background: rgba(255,255,255,.96); border: 1px solid var(--rule); border-radius: .5rem;
  box-shadow: 0 6px 24px rgba(28,28,28,.06);
}
.plan-empty-title { margin: 0; font: 600 1.05rem var(--head); color: var(--ink); }
.plan-empty-hint { margin: 0 0 .5rem; font-size: .9rem; color: var(--muted); }
.plan-empty .chip-clear { margin: 0; background: #fff; }

/* ---- stage ---- */
.plan-stage { position: relative; flex: 1; min-height: 0; overflow: hidden; user-select: none; }
.plan-svg { position: absolute; inset: 0; width: 100%; height: 100%; cursor: grab; touch-action: none; }
.plan-svg.is-dragging { cursor: grabbing; }
.ground { fill: #fff; }

.block-cap { font: 1.25px "Source Sans 3", sans-serif; fill: var(--ink); }
.block-letter { font-family: var(--mono); font-weight: 500; fill: var(--muted); letter-spacing: .06px; }
.axis-no { font: .95px var(--mono); fill: #9aa39e; }

.plot { cursor: pointer; outline: none; }
.plot-fill { stroke: #6b736e; stroke-width: .08; transition: fill-opacity .2s, stroke-width .12s, stroke .12s; }
.plot.is-v3 .plot-fill { fill: var(--v3); } .plot.is-v2 .plot-fill { fill: var(--v2); }
.plot.is-v1 .plot-fill { fill: var(--v1); } .plot.is-v0 .plot-fill { fill: var(--v0); }
.plot.is-buffer { fill: none; stroke: #c4cbc7; stroke-width: .1; stroke-dasharray: .5 .4; }
.plot:hover .plot-fill, .plot:focus-visible .plot-fill { stroke: var(--green); stroke-width: .2; }
.plot.is-selected .plot-fill { stroke: var(--green); stroke-width: .35; }
.plot-id { font: .9px var(--mono); fill: var(--muted); }
.plot-name { font-family: var(--mono); font-weight: 500; fill: var(--ink); }
.plot-by { font-family: "Source Sans 3", sans-serif; fill: var(--muted); }
.plot-cran { fill: var(--green); }
.plot.is-out .plot-fill { fill-opacity: .3; stroke: #c4cbc7; }
.plot.is-out text, .plot.is-out circle, .plot.is-out .crop { opacity: .2; }

/* ---- crops ---- */
.soil { stroke: var(--soil); stroke-width: .08; stroke-linecap: round; }
.stem, .leaf, .awn, .stubble { fill: none; stroke-linecap: round; }
.stem { stroke-width: .1; } .leaf { stroke-width: .09; } .awn { stroke-width: .045; } .stubble { stroke-width: .1; }
.g-v3 .stem, .g-v3 .leaf { stroke: var(--c3); }
.g-v3 .grain { fill: var(--c3-ear); } .g-v3 .awn { stroke: var(--c3-ear); }
.g-v2 .stem, .g-v2 .leaf { stroke: var(--c2); }
.g-v1 .stem, .g-v1 .leaf { stroke: var(--c1); }
.g-v0 .stubble { stroke: var(--c0); }
.plant, .sway { transform-box: fill-box; transform-origin: 50% 100%; }
.plant { animation: grow .9s cubic-bezier(.2,.7,.2,1) both; animation-delay: var(--d, 0s); }
.plot:hover .sway, .plot.is-selected .sway { animation: sway 2.2s ease-in-out infinite alternate; animation-delay: calc(-1 * var(--s, 0s)); }
@keyframes grow { from { transform: scaleY(0); opacity: 0; } to { transform: scaleY(1); opacity: 1; } }
@keyframes sway { from { transform: rotate(-3deg); } to { transform: rotate(3deg); } }
@media (prefers-reduced-motion: reduce) { .plant, .sway { animation: none !important; } }
.plot.is-out { cursor: default; }

/* ---- overlays ---- */
.legend { position: absolute; left: 2rem; bottom: 1.1rem; z-index: 2; }
.legend-toggle { display: none; }
.legend-items {
  display: flex; gap: 1.1rem; flex-wrap: wrap; align-items: center;
  font-size: .78rem; color: var(--muted);
  padding: .3rem .6rem; background: rgba(255,255,255,.92);
}
.legend span { display: inline-flex; align-items: center; gap: .4rem; }
.legend small { opacity: .75; }
.sw { width: .75rem; height: .75rem; display: inline-block; border: 1px solid var(--ink); flex: none; }
.lg-crop { width: 1.1rem; height: 1.9rem; overflow: visible; }
.legend-note { font-style: italic; }
.sw.is-buffer { border: 1px dashed #9aa39e; }
.sw.is-cran { border: 0; width: .45rem; height: .45rem; border-radius: 50%; background: var(--green); margin-inline: .15rem; }

.zoom { position: absolute; right: 2rem; bottom: 1rem; z-index: 2; display: flex; border: 1px solid var(--rule); background: #fff; }
.zoom button { font: 500 .9rem var(--mono); width: 2.2rem; height: 2rem; border: 0; background: none; color: var(--ink); cursor: pointer; }
.zoom button + button { border-left: 1px solid var(--rule); }
.zoom button:hover { background: var(--surface); }
#z-fit { font: .72rem "Source Sans 3", sans-serif; width: auto; padding: 0 .7rem; letter-spacing: .04em; text-transform: uppercase; }

/* ---- entry panel ---- */
.entry {
  position: absolute; top: 0; right: 0; bottom: 0; z-index: 3; width: min(26rem, 100%);
  background: #fff; border-left: 1px solid var(--rule);
  transform: translateX(100%); transition: transform .26s cubic-bezier(.2,.7,.2,1);
  overflow-y: auto; user-select: text;
}
.entry.is-open { transform: none; box-shadow: -12px 0 30px rgba(28,28,28,.06); }
.e-close { position: absolute; top: .7rem; right: .8rem; width: 1.9rem; height: 1.9rem; border: 0; background: none; font-size: 1.3rem; line-height: 1; color: var(--muted); cursor: pointer; }
.e-close:hover { color: var(--ink); }
.e-body { padding: 1.5rem 1.75rem 2rem; }
.e-crop { display: block; width: 100%; height: auto; margin: .2rem 0 1rem; overflow: visible; }
.e-kicker { margin: 0 0 .5rem; font: .72rem var(--mono); color: var(--muted); letter-spacing: .04em; }
.e-name { margin: 0 0 .15rem; font: 500 1.6rem var(--mono); letter-spacing: -.01em; }
.e-by { margin: 0 0 .6rem; color: var(--muted); font-size: .92rem; }
.e-status { display: flex; align-items: center; gap: .3rem; margin: 0 0 1rem; font-size: .88rem; font-weight: 600; }
.e-status span { font-weight: 400; color: var(--muted); }
.e-desc { line-height: 1.55; margin: 0 0 1.1rem; font-size: .95rem; }
.e-list { margin: 0 0 1.2rem; border-top: 1px solid var(--ink); }
.e-row { display: flex; justify-content: space-between; gap: 1rem; padding: .36rem 0; border-bottom: 1px solid var(--rule); }
.e-row dt { font-weight: 400; color: var(--muted); font-size: .84rem; }
.e-row dd { margin: 0; font-family: var(--mono); font-size: .86rem; text-align: right; }
.e-h { margin: .9rem 0 .2rem; font: 600 .66rem var(--head); text-transform: uppercase; letter-spacing: .08em; color: var(--muted); }
.e-tags { margin: 0; font-size: .88rem; line-height: 1.6; }
.e-tags .dot { color: var(--sage); padding: 0 .45rem; }
.e-links { margin: 1.4rem 0 0; display: flex; gap: 1.5rem; font-size: .9rem; }
.e-links a { color: var(--green); font-weight: 600; text-decoration: none; border-bottom: 1px solid var(--sage); }
.e-links a:hover { border-bottom-color: var(--green); }

.visually-hidden { position: absolute !important; width: 1px; height: 1px; margin: -1px; overflow: hidden; clip: rect(0 0 0 0); white-space: nowrap; }

/* ---- small screens --------------------------------------------------------
   Compact chrome for phones in either orientation; keep the two queries in
   step with STACK and SHORT in the script. */
@media (max-width: 700px), (max-height: 520px) {
  .plan-bar { padding: .6rem 1rem .5rem; }
  .plan-title h1 { font-size: 1.15rem; }
  .plan-title p { font-size: .8rem; }
  .plan-search { font-size: 1rem; }   /* 16px+ stops iOS zooming the page on focus */
  .filter-panel { right: 1rem; }
  .plan-active { left: .75rem; right: .75rem; top: .6rem; }

  /* key folds into a button so it doesn't cover the field */
  .legend { left: .75rem; bottom: .75rem; }
  .legend-toggle {
    display: inline-flex; align-items: center; min-height: 2.5rem; padding: 0 .95rem;
    font: inherit; font-size: .85rem; color: var(--ink);
    background: #fff; border: 1px solid var(--rule); border-radius: 999px; cursor: pointer;
  }
  .legend-toggle[aria-expanded="true"] { border-color: var(--sage); background: var(--surface); }
  .legend-items {
    display: none; position: absolute; left: 0; bottom: calc(100% + .5rem);
    flex-direction: column; align-items: flex-start; gap: .35rem; width: max-content;
    max-height: calc(100vh - 10rem); overflow-y: auto;
    padding: .7rem .9rem; font-size: .82rem;
    background: #fff; border: 1px solid var(--rule); border-radius: .5rem; box-shadow: 0 6px 24px rgba(28,28,28,.08);
  }
  .legend.is-open .legend-items { display: flex; }

  .zoom { right: .75rem; bottom: .75rem; border-radius: 999px; overflow: hidden; }
  .zoom button { width: 2.6rem; height: 2.5rem; }
  .e-close { width: 2.75rem; height: 2.75rem; top: .35rem; right: .35rem; z-index: 2; }
}

/* phone portrait: search under the title, entry as a bottom sheet */
@media (max-width: 700px) {
  .plan-head { flex-direction: column; align-items: stretch; gap: .55rem; }
  .plan-tools { flex: 1 1 auto; }
  /* full-width dropdown under the header on phones */
  .filter-panel { left: .75rem; right: .75rem; width: auto; top: calc(100% - .25rem); max-height: calc(100dvh - var(--nav-h, 58px) - 9rem); }
  .fp-show { padding: .65rem 1.4rem; }
  /* the plot stays visible in the strip above the sheet */
  .entry {
    top: auto; left: 0; width: 100%; height: min(66%, 36rem);
    border-left: 0; border-top: 1px solid var(--rule); border-radius: .9rem .9rem 0 0;
    transform: translateY(100%);
  }
  .entry.is-open { box-shadow: 0 -12px 30px rgba(28,28,28,.08); }
  .entry::before {
    content: ""; display: block; position: sticky; top: 0; height: 1.1rem; z-index: 1;
    background: linear-gradient(#fff 70%, transparent);
  }
  .entry::after {
    content: ""; position: absolute; top: .45rem; left: 50%; width: 2.5rem; height: .25rem;
    margin-left: -1.25rem; border-radius: 999px; background: var(--rule);
  }
  .e-body { padding: .4rem 1.25rem 2rem; }
  .e-name { font-size: 1.35rem; }
  .e-crop { margin-bottom: .75rem; }
}

/* phone landscape: every pixel of height goes to the field */
@media (max-height: 520px) and (min-width: 701px) {
  .plan-bar { padding: .45rem 1rem .4rem; }
  .plan-title { display: flex; align-items: baseline; gap: .7rem; }
  .plan-title p { margin: 0; }
  .plan-search { padding: .35rem .9rem; }
  .filter-btn { height: 2.2rem; }
  .filter-panel { max-height: calc(100dvh - var(--nav-h, 58px) - 4rem); }
  .entry { width: min(24rem, 50%); }
  .e-body { padding: 1rem 1.25rem 1.5rem; }
}
</style>

<script>
(function () {
  const plan  = document.getElementById('plan');
  const stage = plan.querySelector('.plan-stage');
  const svg   = document.getElementById('plan-svg');
  const world = document.getElementById('world');
  const ground = svg.querySelector('.ground');
  const blocks = Array.from(svg.querySelectorAll('.block'));
  const D = svg.dataset;
  const PW = +D.pw, PH = +D.ph, CG = +D.cg, RG = +D.rg, BG = +D.bg, PAD = +D.pad;
  const NS = 'http://www.w3.org/2000/svg';
  // Keep in step with the small-screen @media rules.
  const STACK = matchMedia('(max-width: 700px)');   // phone portrait
  const SHORT = matchMedia('(max-height: 520px)');  // phone landscape
  const currentMode = () => STACK.matches ? 'stack' : SHORT.matches ? 'row' : 'grid';
  let FW = 1, FH = 1, layoutMode = null;

  const svgEl = (tag, attrs, text) => {
    const e = document.createElementNS(NS, tag);
    for (const a in attrs) e.setAttribute(a, attrs[a]);
    if (text != null) e.textContent = text;
    return e;
  };

  // ---- layout -------------------------------------------------------------
  //   grid  (desktop/tablet): blocks side by side, each roughly square; the
  //         whole plan fits on screen.
  //   stack (phone portrait): blocks stacked, one plot per range; fits the
  //         width and you scroll down through it.
  //   row   (phone landscape): every block a single range, side by side;
  //         fits the height and you scroll across it.
  function layoutPlan(mode) {
    const narrow = mode === 'stack';
    // Stacked, the plan is zoomed to the screen width, so the usual margin
    // above it would become a large blank band; start close to the top.
    let x0 = PAD, y0 = narrow ? 1.5 : PAD, maxR = 0, maxB = 0;
    const capY = 3.4;  // room above a block for its caption
    blocks.forEach(b => {
      const ps = Array.from(b.querySelectorAll('.plot[data-id]'));
      const n = ps.length;
      const ncol = mode === 'stack' ? 1 : mode === 'row' ? n : Math.max(1, Math.ceil(Math.sqrt(n * PH / PW)));
      const nrow = Math.ceil(n / ncol);
      const w = ncol * PW + (ncol - 1) * CG, h = nrow * PH + (nrow - 1) * RG;

      // Caption: letter and organisation on one line when wide; on phones the
      // organisation wraps onto lines under the letter to fit the plot width.
      const cap = b.querySelector('.block-cap');
      cap.querySelectorAll('.block-name').forEach(t => t.remove());
      const name = cap.dataset.name;
      let lines = [name];
      if (narrow) {
        lines = []; let line = '';
        const maxChars = Math.floor(PW / 0.62);
        name.split(/\s+/).forEach(word => {
          if (line && (line + ' ' + word).length > maxChars) { lines.push(line); line = word; }
          else line = line ? line + ' ' + word : word;
        });
        if (line) lines.push(line);
      }
      lines.forEach((ln, i) => cap.appendChild(svgEl('tspan', narrow
        ? { class: 'block-name', x: 0, dy: i === 0 ? 1.55 : 1.45 }
        : { class: 'block-name', dx: 1.6 }, ln)));
      const capH = narrow ? 2.4 + lines.length * 1.5 : capY;

      const bx = narrow ? PAD : x0;
      const by = narrow ? y0 + capH : PAD + capY;
      cap.setAttribute('x', bx); cap.setAttribute('y', by - capH + 1.1);
      cap.querySelectorAll('tspan.block-name[x]').forEach(t => t.setAttribute('x', bx));
      const capW = cap.getComputedTextLength ? (narrow ? Math.min(cap.getComputedTextLength(), PW) : cap.getComputedTextLength()) : 0;

      ps.forEach((p, i) => {
        const r = Math.floor(i / ncol), c = i % ncol;
        const x = bx + c * (PW + CG), y = by + r * (PH + RG);
        p.setAttribute('transform', `translate(${x} ${y})`);
        Object.assign(p.dataset, { cx: x + PW / 2, cy: y + PH / 2, row: r + 1, col: c + 1 });
      });

      // Buffers and range/column numbers depend on the layout, so rebuild them.
      const extras = b.querySelector('.block-extras');
      extras.replaceChildren();
      for (let i = n; i < ncol * nrow; i++) {
        const r = Math.floor(i / ncol), c = i % ncol;
        extras.appendChild(svgEl('rect', { class: 'plot is-buffer', x: bx + c * (PW + CG), y: by + r * (PH + RG), width: PW, height: PH }));
      }
      for (let r = 0; r < nrow; r++)
        extras.appendChild(svgEl('text', { class: 'axis-no', x: bx - 1, y: by + r * (PH + RG) + PH / 2 + 0.45, 'text-anchor': 'end' }, r + 1));
      for (let c = 0; c < ncol; c++)
        extras.appendChild(svgEl('text', { class: 'axis-no', x: bx + c * (PW + CG) + PW / 2, y: by + h + 2.2, 'text-anchor': 'middle' }, c + 1));

      maxR = Math.max(maxR, bx + Math.max(w, capW));
      maxB = Math.max(maxB, by + h + 3);
      if (narrow) y0 = by + h + 3 + BG * 0.6;
      else x0 += Math.max(w + BG, capW + 4);
    });
    FW = maxR + PAD; FH = maxB + PAD;
    ground.setAttribute('width', FW); ground.setAttribute('height', FH);
  }

  // ---- pan / zoom ---------------------------------------------------------
  let k = 1, tx = 0, ty = 0, anim = null;
  const render = () => world.setAttribute('transform', `translate(${tx} ${ty}) scale(${k})`);
  // grid: fit the whole plan. stack: fit the width, start at the top.
  // row: fit the height, start at the left.
  function fitParams() {
    const W = svg.clientWidth, H = svg.clientHeight;
    if (layoutMode === 'stack') {
      const pad = 12, kk = (W - pad * 2) / FW;
      return { k: kk, tx: (W - FW * kk) / 2, ty: FH * kk <= H - pad * 2 ? (H - FH * kk) / 2 : pad };
    }
    if (layoutMode === 'row') {
      const pad = 8, kk = (H - pad * 2) / FH;
      return { k: kk, ty: (H - FH * kk) / 2, tx: FW * kk <= W - pad * 2 ? (W - FW * kk) / 2 : pad };
    }
    const pad = 48, kk = Math.min((W - pad * 2) / FW, (H - pad * 2) / FH);
    return { k: kk, tx: (W - FW * kk) / 2, ty: (H - FH * kk) / 2 };
  }
  const clampK = v => Math.max(fitParams().k * 0.6, Math.min(v, 60));
  function stopAnim() { if (anim) { cancelAnimationFrame(anim); anim = null; } }
  function animateTo(t, ms) {
    stopAnim();
    const f = { k, tx, ty }, t0 = performance.now(), ease = x => 1 - Math.pow(1 - x, 3);
    (function step(now) {
      const p = ease(Math.min(1, (now - t0) / ms));
      k = f.k + (t.k - f.k) * p; tx = f.tx + (t.tx - f.tx) * p; ty = f.ty + (t.ty - f.ty) * p;
      render(); anim = p < 1 ? requestAnimationFrame(step) : null;
    })(t0);
  }
  function fit(animate = true) {
    const p = fitParams();
    animate ? animateTo(p, 320) : (stopAnim(), ({ k, tx, ty } = p), render());
  }
  function zoomAt(f, cx, cy) {
    stopAnim();
    const nk = clampK(k * f);
    tx = cx - (cx - tx) * (nk / k); ty = cy - (cy - ty) * (nk / k); k = nk; render();
  }
  svg.addEventListener('wheel', e => {
    e.preventDefault();
    const r = svg.getBoundingClientRect();
    zoomAt(Math.exp(-e.deltaY * 0.0015), e.clientX - r.left, e.clientY - r.top);
  }, { passive: false });

  // One pointer pans (or taps); two pointers pinch-zoom around their midpoint.
  const pts = new Map();
  let drag = null, pinch = null;
  const local = (x, y) => { const r = svg.getBoundingClientRect(); return [x - r.left, y - r.top]; };
  svg.addEventListener('pointerdown', e => {
    if (e.pointerType === 'mouse' && e.button !== 0) return;
    stopAnim();
    pts.set(e.pointerId, { x: e.clientX, y: e.clientY });
    try { svg.setPointerCapture(e.pointerId); } catch (_) {}
    if (pts.size === 1) {
      // Remember the plot now: once the pointer is captured, later events
      // are retargeted to the svg and the plot is no longer e.target.
      drag = { x: e.clientX, y: e.clientY, tx, ty, moved: false, plot: e.target.closest('.plot[data-id]') };
    } else if (pts.size === 2) {
      drag = null;  // a second finger makes this a pinch, never a tap
      const [a, b] = [...pts.values()];
      const [mx, my] = local((a.x + b.x) / 2, (a.y + b.y) / 2);
      pinch = { d0: Math.hypot(b.x - a.x, b.y - a.y) || 1, k0: k, tx0: tx, ty0: ty, mx0: mx, my0: my };
    }
  });
  svg.addEventListener('pointermove', e => {
    if (!pts.has(e.pointerId)) return;
    pts.set(e.pointerId, { x: e.clientX, y: e.clientY });
    if (pinch && pts.size >= 2) {
      const [a, b] = [...pts.values()];
      const [mx, my] = local((a.x + b.x) / 2, (a.y + b.y) / 2);
      const nk = clampK(pinch.k0 * Math.hypot(b.x - a.x, b.y - a.y) / pinch.d0);
      // keep the point that started under the fingers under the fingers
      tx = mx - (pinch.mx0 - pinch.tx0) * (nk / pinch.k0);
      ty = my - (pinch.my0 - pinch.ty0) * (nk / pinch.k0);
      k = nk; render();
      return;
    }
    if (!drag) return;
    const dx = e.clientX - drag.x, dy = e.clientY - drag.y;
    if (Math.hypot(dx, dy) > 4) { drag.moved = true; svg.classList.add('is-dragging'); }
    if (drag.moved) { tx = drag.tx + dx; ty = drag.ty + dy; render(); }
  });
  function endPointer(e) {
    if (!pts.has(e.pointerId)) return;
    pts.delete(e.pointerId);
    if (pinch) { if (pts.size < 2) pinch = null; return; }  // lifting after a pinch is not a tap
    if (!drag || pts.size) return;
    const { moved, plot } = drag; drag = null; svg.classList.remove('is-dragging');
    if (moved || e.type === 'pointercancel') return;
    // Fall back to a hit test at the release point, which is immune to
    // pointer-capture retargeting.
    const g = plot || document.elementFromPoint(e.clientX, e.clientY)?.closest('.plot[data-id]');
    if (g && !g.classList.contains('is-out')) open(g); else close();
  }
  svg.addEventListener('pointerup', endPointer);
  svg.addEventListener('pointercancel', endPointer);

  const mid = () => [svg.clientWidth / 2, svg.clientHeight / 2];
  document.getElementById('z-in').onclick  = () => zoomAt(1.5, ...mid());
  document.getElementById('z-out').onclick = () => zoomAt(1 / 1.5, ...mid());
  document.getElementById('z-fit').onclick = () => { close(); fit(); };

  // ---- key (collapsible on phones) ----------------------------------------
  const legend = document.getElementById('legend');
  const legendBtn = legend.querySelector('.legend-toggle');
  legendBtn.addEventListener('click', () => {
    const open = !legend.classList.contains('is-open');
    legend.classList.toggle('is-open', open);
    legendBtn.setAttribute('aria-expanded', String(open));
  });

  // ---- entry panel ----
  const panel = document.getElementById('entry');
  const body  = document.getElementById('e-body');
  const plots = Array.from(svg.querySelectorAll('.plot[data-id]'));
  let selected = null;

  function open(g) {
    plots.forEach(p => p.classList.toggle('is-selected', p === g));
    selected = g;
    body.replaceChildren(document.querySelector(`template[data-for="${g.dataset.id}"]`).content.cloneNode(true));
    body.querySelector('.e-pos').textContent = `range ${g.dataset.row}, column ${g.dataset.col}`;
    panel.classList.add('is-open'); panel.setAttribute('aria-hidden', 'false');
    legend.classList.remove('is-open'); legendBtn.setAttribute('aria-expanded', 'false');
    body.parentElement.scrollTop = 0;
    animateTo(frame(g), 380);
  }
  // Where to put the view so plot g sits centred in the space the panel
  // leaves free: to its left on wide screens, above the sheet on phones.
  function frame(g) {
    const W = svg.clientWidth, H = svg.clientHeight;
    const sheet = STACK.matches;
    const fw = sheet ? W : W - Math.min(panel.offsetWidth, W);
    const fh = sheet ? H - Math.min(panel.offsetHeight, H) : H;
    // Wide: about a third of the free space, so neighbours stay in view.
    // Phone: as large as the free strip above the sheet comfortably allows.
    const nk = sheet ? Math.min(fw * 0.8 / PW, fh * 0.75 / PH, 40)
                     : Math.max(k, Math.min(fw * 0.3 / PW, fh * 0.3 / PH, 30));
    return { k: nk, tx: fw / 2 - g.dataset.cx * nk, ty: fh / 2 - g.dataset.cy * nk };
  }
  function close() {
    if (!selected) return;
    selected.classList.remove('is-selected'); selected = null;
    panel.classList.remove('is-open'); panel.setAttribute('aria-hidden', 'true');
  }
  document.getElementById('e-close').onclick = close;
  addEventListener('keydown', e => {
    if (e.key === 'Escape') close();
    if ((e.key === 'Enter' || e.key === ' ') && document.activeElement?.classList.contains('plot')) {
      e.preventDefault(); open(document.activeElement);
    }
  });

  // ---- search + faceted filters ----
  // Within a group, chips are OR (any selected keyword); across groups and
  // search, AND. Each chip shows how many plots you'd see if you added it.
  const input   = document.getElementById('pl-search');
  const chips   = Array.from(document.querySelectorAll('.facet-chip'));
  const clearEl = document.getElementById('pl-clear');
  const emptyEl = document.getElementById('pl-empty');
  const nMatch  = document.getElementById('n-match');
  const active  = { kw: new Set(), status: new Set(), cran: new Set() };
  const facts = new Map(plots.map(p => [p, {
    search: p.dataset.search,
    kw: (p.dataset.keywords || '').split('|').filter(Boolean),
    status: [p.dataset.status],
    cran: p.dataset.cran === '1' ? ['1'] : []
  }]));

  function matches(p, term, skip) {
    const f = facts.get(p);
    if (term && !f.search.includes(term)) return false;
    for (const g of Object.keys(active)) {
      if (g === skip || active[g].size === 0) continue;
      if (!f[g].some(v => active[g].has(v))) return false;
    }
    return true;
  }

  const fBtn    = document.getElementById('pl-filter-btn');
  const fPanel  = document.getElementById('pl-filters');
  const fBadge  = document.getElementById('pl-filter-n');
  const fShow   = document.getElementById('pl-show');
  const activeEl = document.getElementById('pl-active');

  function apply() {
    const term = input.value.trim().toLowerCase();
    let n = 0;
    plots.forEach(p => {
      const ok = matches(p, term);
      p.classList.toggle('is-out', !ok);
      p.setAttribute('tabindex', ok ? '0' : '-1');
      if (ok) n++;
    });
    nMatch.textContent = n;

    chips.forEach(c => {
      const g = c.dataset.group, k = c.dataset.key;
      const count = plots.filter(p => matches(p, term, g) && facts.get(p)[g].includes(k)).length;
      const on = active[g].has(k);
      c.querySelector('.chip-n').textContent = count;
      c.disabled = count === 0 && !on;
      c.setAttribute('aria-label', `${c.querySelector('.chip-label').textContent}, ${count} plot${count === 1 ? '' : 's'}`);
    });

    const nOn = Object.values(active).reduce((a, s) => a + s.size, 0);
    const filtering = !!term || nOn > 0;
    clearEl.hidden = !filtering;
    fBadge.hidden = nOn === 0;
    fBadge.textContent = nOn;
    fBtn.classList.toggle('is-active', nOn > 0);
    fBtn.setAttribute('aria-label', nOn ? `Filters, ${nOn} active` : 'Filters');
    fShow.textContent = n === 0 ? 'No matches' : `Show ${n} plot${n === 1 ? '' : 's'}`;
    fShow.disabled = n === 0;
    renderActive();
    emptyEl.hidden = n !== 0;
    if (selected && selected.classList.contains('is-out')) close();
    syncUrl(term);
  }

  // Keep the filter state in the URL so a filtered view can be shared.
  // (Not ?q=: Quarto's site search claims and strips that parameter.)
  function syncUrl(term) {
    const u = new URL(location.href);
    ['search', 'k', 'status', 'cran'].forEach(x => u.searchParams.delete(x));
    if (term) u.searchParams.set('search', term);
    if (active.kw.size) u.searchParams.set('k', [...active.kw].join(','));
    if (active.status.size) u.searchParams.set('status', [...active.status].join(','));
    if (active.cran.size) u.searchParams.set('cran', '1');
    history.replaceState(null, '', u);
  }

  // Removable pills for whatever is switched on, so the filters stay
  // visible while the panel is closed.
  function renderActive() {
    const on = chips.filter(c => c.getAttribute('aria-pressed') === 'true');
    activeEl.replaceChildren();
    on.forEach(c => {
      const label = c.querySelector('.chip-label').textContent;
      const b = document.createElement('button');
      b.type = 'button'; b.className = 'chip active-pill';
      b.setAttribute('aria-label', `Remove filter: ${label}`);
      b.append(label);
      const x = document.createElement('span'); x.className = 'x'; x.setAttribute('aria-hidden', 'true'); x.textContent = '×';
      b.append(x);
      b.addEventListener('click', () => { setChip(c, false); apply(); });
      activeEl.append(b);
    });
    if (on.length > 1) {
      const b = document.createElement('button');
      b.type = 'button'; b.className = 'active-clear'; b.textContent = 'Clear all';
      b.addEventListener('click', clearAll);
      activeEl.append(b);
    }
    activeEl.hidden = on.length === 0;
  }

  function setFilters(open, returnFocus = true) {
    fPanel.hidden = !open;
    fBtn.setAttribute('aria-expanded', String(open));
    if (open) {
      legend.classList.remove('is-open'); legendBtn.setAttribute('aria-expanded', 'false');
      (fPanel.querySelector('.facet-chip:not(:disabled)') || fPanel.querySelector('.fp-close')).focus({ preventScroll: true });
    } else if (returnFocus) fBtn.focus({ preventScroll: true });
  }
  fBtn.addEventListener('click', () => setFilters(fPanel.hidden));
  document.getElementById('pl-filters-close').addEventListener('click', () => setFilters(false));
  fShow.addEventListener('click', () => setFilters(false));
  // Click or tap anywhere else closes it (without stealing focus back).
  document.addEventListener('pointerdown', e => {
    if (!fPanel.hidden && !fPanel.contains(e.target) && !fBtn.contains(e.target)) setFilters(false, false);
  }, true);
  // Escape closes the panel first, before anything else on the page reacts.
  document.addEventListener('keydown', e => {
    if (e.key === 'Escape' && !fPanel.hidden) { e.stopPropagation(); setFilters(false); }
  }, true);

  function setChip(c, on) {
    active[c.dataset.group][on ? 'add' : 'delete'](c.dataset.key);
    c.setAttribute('aria-pressed', String(on));
  }
  function clearAll() {
    chips.forEach(c => setChip(c, false));
    input.value = '';
    apply();
  }

  input.addEventListener('input', apply);
  input.addEventListener('keydown', e => { if (e.key === 'Escape' && input.value) { e.stopPropagation(); input.value = ''; apply(); } });
  chips.forEach(c => c.addEventListener('click', () => {
    setChip(c, c.getAttribute('aria-pressed') !== 'true'); apply();
  }));
  clearEl.addEventListener('click', clearAll);
  document.getElementById('pl-empty-clear').addEventListener('click', clearAll);

  // Restore state from the URL. ?k= matches the database page's deep links.
  const params = new URLSearchParams(location.search);
  input.value = params.get('search') || params.get('q') || '';
  const want = { kw: params.get('k'), status: params.get('status'), cran: params.get('cran') };
  chips.forEach(c => {
    const list = (want[c.dataset.group] || '').toLowerCase().split(',').map(x => x.trim());
    if (list.includes(c.dataset.key)) setChip(c, true);
  });

  apply();

  function sizeToNav() {
    const nav = document.querySelector('#quarto-header') || document.querySelector('.navbar');
    plan.style.setProperty('--nav-h', (nav ? nav.getBoundingClientRect().bottom : 0) + 'px');
  }
  // Re-lay out only when crossing the phone breakpoint; otherwise just keep
  // the view sensible for the new size.
  function relayout() {
    sizeToNav();
    const mode = currentMode();
    if (mode !== layoutMode) { layoutMode = mode; layoutPlan(mode); }
    if (selected) {
      const pos = body.querySelector('.e-pos');
      if (pos) pos.textContent = `range ${selected.dataset.row}, column ${selected.dataset.col}`;
      stopAnim(); ({ k, tx, ty } = frame(selected)); render();
    }
    else fit(false);
  }
  relayout();
  let pending = null;
  const ro = new ResizeObserver(() => { cancelAnimationFrame(pending); pending = requestAnimationFrame(relayout); });
  ro.observe(stage);
  const header = document.querySelector('#quarto-header');
  if (header) ro.observe(header);
  if (document.fonts) document.fonts.ready.then(() => { layoutMode = null; relayout(); });
})();
</script>
```
